Freezing of a Weeks-Chandler-Anderson Fluid using classical Density Functional Theory

### 2020-10-17 papers/thesis-kirstie/thesis.tex
more updates

Classically, fluids have been modeled by a collection of small, impenetrable spheres colliding 
with one another like billiard balls. In the simplest case, the spheres are treated as having 
only a replulsive potential. This model, called the Hard-Sphere fluid, has served as the primary 
reference system for fluids. This is due to the fact that the repulsion between atoms is much 
stronger than their attraction, and so the repulsive potential, which is descriptive of 
hard spheres, dominates the interactions between atoms. 
In 1954 Zwanzig introduced the Square-well fluid which is a perturbation of the Hard-Sphere fluid 
created by adding a hard-edge attractive potential to the repulsive potential of the Hard-Sphere 
fluid~\cite{ZwanzigSqrWell}. In 1964 Rowlinson studied spheres with softly fading replusive potentials which he 
modeled by making the radius of the spheres temperature dependent~\cite{rowlinson1964statistical}. 
This reflects the idea that soft-spheres penetrate more deeply into each other 
when they collide at higher temperatures. 
%At higher temperatures the spheres would penetrate more deeply into each other when they collide, 
%and thus be represented by spheres with smaller radii. 
In 1967 Barker and Henderson combined the perturabtion approach of Zwanzig 
with Rowlinson's idea of using temperature-dependent radii to model spheres with 
soft repulsive potentials to create the Barker-Henderson model fluid~\cite{barker1967perturbation}, 
where a soft-sphere fluid by is modeled by using a Hard-Sphere fluid with hard spheres of 
larger radii at low temperatures, and smaller radii at high temperatures. 
%
%an expression for the excess Helmholtz free energy of inhomogeneous fluids.
% ?  Percus-Yevick applied Density Functional Theory (DFT) to find an expression for the excess 
%Helmholtz free energy of an inhomogeneous Hard-Sphere fluid. 
%
%In 1967 Barker and Henderson~\cite{barker1967perturbation} adapted the perturbation method of Zwanzig 
%to use with fluids composed of spheres with softly fading potentials rather than the hard edges of the spheres 
%in the Hard-Sphere fluid. 
%This was done by making the radius of the spheres temperature dependent, 
%so that at low temperatures the Soft-Sphere fluid could be modeled by hard spheres with large radii, 
%and at high temperatures the Soft-Sphere fluid could be modeled by hard spheres with smaller radii. 
%
%A larger radii cooresponds to a deeper level of penetration of two soft spheres into each other 
%when they collide at higher temperatures.
%The different size radii coorespond to the different depths to which two soft 
%spheres would penetrate into each other when they collide. 
%This reflects the idea that soft-spheres penetrate more deeply into each other when 
%they collide at higher temperatures. 
In 1971 Weeks, Chandler, and Anderson used temperature and density dependent hard-sphere 
radii to model spheres with the more realistic softly fading repulsive portion of the 
Leonard-Jones potential~\cite{andersen1971relationship}, which became known as the 
Weeks-Chandler-Anderson model fluid.

While models for particular fluids were being created, powerful methods to describe the properties 
of fluids in general were also being developed. In the late 1950s, Scaled Particle 
Theory (SPT)~\cite{ReissSPT} was introduced from which an equation of state, and
an expression for the excess Helmholtz free energy of a homogeneous Hard-Sphere fluid can be 
obtained~\cite{Hansen}. \color{red}ADD Percus-Yevick introduced an alternative method...OZ... \color{black}
%Percus and Yevick took another approach to the Hard-Sphere fluid.
Led by the consistency bewteen results from SPT and Percus-Yevick, in 1961 Morita and Hiroike 
introduced a density functional approach~\cite{MoritaDFT} to finding the equilibrium grand free energy 
(the equilibrium Helmholtz free energy in the case of a canonical ensemble), and the cooresponding equilibrium 
number density profile for inhomogeneous fluids. To use this approach, a functional must be 
constructed for a particular fluid.
%DFT Evans 1979
In 1989 Rosenfeld introduced Fundamental Measure Theory (FMT)~\cite{rosenfeld1989} 
(which can be constructed from SPT~\cite{santos2012phi3}) to form a functional for 
%which provides an expression for the excess Helmholtz free energy of an inhomogeneous 
the Hard-Sphere fluid. 
Tarazona later improved on the method by introducing a tensor version of the 
functional~\cite{tarazonaphi3}, and in 1999 Schmidt adopted FMT for soft spheres 
forming Soft Fundamental Measure Theory (SFMT)\cite{schmidt1999density}.

Schmidt applied his Soft Fundamental Measure Theory to various fluids with soft repulsive 
and soft attractive potentials, and developed an erf model for 
spheres with a steep, short-range soft repulsive potential~\cite{schmidt2000fluid}. In this paper, 
Density Functional Theory (DFT) will be used to predict the freezing of a 
Weeks-Chandler-Anderson model fluid by applying SFMT with Schmidt's proposed 
erf model to form the functional for the excess Helmholtz free energy. 
%for generating the weight funcions, and Tarazona's tensor functional. 
Results will be compared to those generated from Monte-Carlo simulations.
%\[{}\]
\newline
\newline
\noindent IIA. Model Fluids and the Weeks-Chandler-Anderson Fluid (WCA)

@@

Phase diagrams for the WCA fluid are shown below. In all cases $\sigma=1$ and $\epsilon=1$.

\begin{figure}
 \centering
 %\includegraphics[width=\columnwidth]{figs/Phase_Diagram_of_T_vs_n}
\caption{Phase Diagram of the reduced termperature verses the reduced density. Green lines indicate the region of coexistence from Monte-Carlo simulations.}
\label{fig:Phase_Diagram_of_T_vs_n}
\end{figure}

\begin{figure}
 \centering
 %\includegraphics[width=\columnwidth]{figs/Phase_Diagram_of_P_vs_T}
\caption{Phase Diagram of the reduced pressure verses the reduced temperature. The green line marks the boundary from Monte-Carlo simulations.}
\label{fig:Phase_Diagram_P_vs_T}
\end{figure}

\begin{figure}
 \begin{center}
  %\includegraphics[width=\columnwidth]{figs/T-vs-n_at_fixed_P}
 \end{center}
\caption{Phase Diagram of the reduced temperature verses the reduced density. The grey region indicates coexistence.}
\label{fig:T-vs-n_at_fixed_P}
\end{figure}

\begin{figure}
 \begin{center}
  %\includegraphics[width=\columnwidth]{figs/p-vs-T_at_fixed_n}
 \end{center}
\caption{Phase Diagram of the reduced pressure verses the reduced temperature. The lines bending downward indicate freezing.}
\label{fig:p-vs-T_at_fixed_density}
\end{figure}

\begin{figure}
 \begin{center}
  %\includegraphics[width=\columnwidth]{figs/p-vs-V_at_fixed_T}
 \end{center}
\caption{Phase Diagram of the reduced pressure verses the reduced volume. The grey region indicates coexistence.}
\label{fig:p-vs-V_at_fixed_T}
\end{figure}

\begin{figure}
 \begin{center}
  %\includegraphics[width=\columnwidth]{figs/p-vs-n_at_fixed_T}
 \end{center}
\caption{Phase Diagram of the reduced pressure verses the reduced density. The grey region indicates coexistence.}
\label{fig:p-vs-n_at_fixed_T}
\end{figure}

@@

\cite{rosenfeld1989}
\cite{tarazona2000}
\cite{schmidt1999density}
\cite{schmidt2000fluid}

\cite{santos2012phi3}

\cite{TschoppBHextended}

%1-21
\cite{lucretius}
\cite{rowlinson1964statistical}
\cite{gil-villegas-1997-SAFT-VR, clark2006developing,
  lafitte2013accurate}
\cite{carnahan1969equation}
\cite{rosenfeld1989, rosenfeld1997,
  roth2002whitebear}
\cite{rosenfeld1989}
%\cite{cuesta1997dimensional, hansen2009fundamental, marechal2013density, hughes2013classical, krebs2014improved}
\cite{cuesta1997dimensional, hansen2009fundamental, marechal2013density}
%\cite{schmidt1999density}
\cite{schmidt2000density, groh2001density, kim2001adsorption,
  sweatman2002fundamental}
%\cite{schmidt2000fluid, sweatman2002fundamental}
\cite{weeks1971}
%\cite{rosenfeld1989}
%\cite{schmidt1999density}
%%\cite{tarazonaFIXME,santos2012phi3}
%\cite{schmidt1999density} 
%\cite{schmidt2000fluid}
%\cite{schmidt1999density, schmidt2000fluid}
%\cite{schmidt1999density}
%\citenum{schmidt2000fluid}
%\citenum{schmidt2000fluid}
%a Face Centered Cubic (FCC) lattice is formed~\cite{reference!}
\cite{barker1967perturbation}
\cite{roth2002whitebear} 
%\cite{roth2002whitebear}
\cite{mikolaj2004structure, eggert2002quantitative,
yarnell1973structure}
\cite{verlet1967computer}

>>>>

Classically, fluids have been modeled by a collection of small, impenetrable spheres colliding 
with one another like billiard balls. In the simplest case, the spheres are treated as having 
only a replulsive potential. This model, called the Hard-Sphere fluid, has served as the primary 
reference system for fluids. This is due to the fact that the repulsion between atoms is much 
stronger than their attraction, and so the repulsive potential, which is descriptive of 
hard spheres, dominates the interactions between atoms. 
In 1954 Zwanzig introduced the Square-well fluid which is a perturbation of the Hard-Sphere fluid 
created by adding a hard-edge attractive potential to the repulsive potential of the Hard-Sphere 
fluid~\cite{ZwanzigSqrWell}. In 1964 Rowlinson studied spheres with softly fading replusive potentials which he 
modeled by making the radius of the spheres temperature dependent~\cite{rowlinson1964statistical}. 
This reflects the idea that soft-spheres penetrate more deeply into each other 
when they collide at higher temperatures. 
%At higher temperatures the spheres would penetrate more deeply into each other when they collide, 
%and thus be represented by spheres with smaller radii. 
In 1967 Barker and Henderson combined the perturabtion approach of Zwanzig 
with Rowlinson's idea of using temperature-dependent radii to model spheres with 
soft repulsive potentials to create the Barker-Henderson model fluid~\cite{barker1967perturbation}, 
where a soft-sphere fluid by is modeled by using a Hard-Sphere fluid with hard spheres of 
larger radii at low temperatures, and smaller radii at high temperatures. 
%
%an expression for the excess Helmholtz free energy of inhomogeneous fluids.
% ?  Percus-Yevick applied Density Functional Theory (DFT) to find an expression for the excess 
%Helmholtz free energy of an inhomogeneous Hard-Sphere fluid. 
%
%In 1967 Barker and Henderson~\cite{barker1967perturbation} adapted the perturbation method of Zwanzig 
%to use with fluids composed of spheres with softly fading potentials rather than the hard edges of the spheres 
%in the Hard-Sphere fluid. 
%This was done by making the radius of the spheres temperature dependent, 
%so that at low temperatures the Soft-Sphere fluid could be modeled by hard spheres with large radii, 
%and at high temperatures the Soft-Sphere fluid could be modeled by hard spheres with smaller radii. 
%
%A larger radii cooresponds to a deeper level of penetration of two soft spheres into each other 
%when they collide at higher temperatures.
%The different size radii coorespond to the different depths to which two soft 
%spheres would penetrate into each other when they collide. 
%This reflects the idea that soft-spheres penetrate more deeply into each other when 
%they collide at higher temperatures. 
In 1971 Weeks, Chandler, and Anderson used temperature and density dependent hard-sphere 
radii to model spheres with the more realistic softly fading repulsive portion of the 
Leonard-Jones potential~\cite{andersen1971relationship}, which is known as the 
Weeks-Chandler-Anderson model fluid.

While models for particular fluids were being created, powerful methods to describe the properties 
of fluids in general were also being developed. In the late 1950s, Scaled Particle 
Theory (SPT)~\cite{ReissSPT} was introduced from which an equation of state, and
an expression for the excess Helmholtz free energy of a homogeneous Hard-Sphere fluid can be 
obtained~\cite{Hansen}. \color{red}ADD Percus-Yevick introduced an alternative method...OZ... \color{black}
%Percus and Yevick took another approach to the Hard-Sphere fluid.
Led by the consistency bewteen results from SPT and Percus-Yevick, in 1961 Morita and Hiroike 
introduced a density functional approach~\cite{MoritaDFT} to finding the equilibrium grand free energy 
(the equilibrium Helmholtz free energy in the case of a canonical ensemble), and the cooresponding equilibrium 
number density profile for inhomogeneous fluids. To use this approach, a functional must be 
constructed for a particular fluid.
%DFT Evans 1979
In 1989 Rosenfeld introduced Fundamental Measure Theory (FMT)~\cite{rosenfeld1989} 
(which can be constructed from SPT~\cite{santos2012phi3}) to form a functional for 
%which provides an expression for the excess Helmholtz free energy of an inhomogeneous 
the Hard-Sphere fluid. 
Tarazona later improved on the method by introducing a tensor version of the 
functional~\cite{tarazonaphi3, santos2012phi3}, and in 1999 Schmidt adopted FMT for soft spheres 
forming Soft Fundamental Measure Theory (SFMT)\cite{schmidt1999density}.

Schmidt applied his Soft Fundamental Measure Theory to various fluids with soft repulsive 
and soft attractive potentials, and developed an erf model for 
spheres with steep, short-range soft repulsive potentials~\cite{schmidt2000fluid}. In this paper, 
Density Functional Theory (DFT) will be used to predict the freezing of a 
Weeks-Chandler-Anderson model fluid by applying SFMT with Schmidt's proposed 
erf model to form the functional for the excess Helmholtz free energy. 
%for generating the weight funcions, and Tarazona's tensor functional. 
Results will be compared to those generated from Monte-Carlo simulations.
%\[{}\]
\newline
\newline
\noindent IIA. Model Fluids and the Weeks-Chandler-Anderson Fluid (WCA)

@@

Phase diagrams for the WCA fluid are shown in Figures \ref{fig:Phase_Diagram_of_T_vs_n} 
and \ref{fig:Phase_Diagram_P_vs_T}. In all cases $\sigma=1$ and $\epsilon=1$. 
In Figure \ref{fig:Phase_Diagram_of_T_vs_n} the reduced temperature $T^*=\frac{kT}{\epsilon}$ 
is plotted against the reduced number density $n^*=n\sigma^3$ for temperatures of kT = 0.5 to 3.0. 
The red area is fluid, and the blue area is crystal. The gray area that separates the fluid, 
and the crystal regions is where some fluid and crystal exist together at varying porportions. 
The region of coexistence slants slightly to the right toward higher density as the temperature 
increases. This is expected since at higher temperatures, the fluid would have to be under 
greater compression before forming a solid. The white lines show results from Monte-Carlo 
simulations of the WCA fluid. Overall, the general shape, and position of the lines correspond
fairly closely to the edges of the region of coexistance predicted from the DFT method, 
but the width of the coexistance region is about one-third larger for the Monte-Carlo results 
for most temperatures. The white line on the right, representing the density of the solid at 
each temperature, is in fairly good aggreement with our DFT results for all temperatures. 
However, the white line on the left, representing the density of the liquid at each temperature, 
is consistenty farther left than the liquid transition densities our method predicted.

Figure \ref{fig:Phase_Diagram_P_vs_T} shows the reduced pressure $p^*=\frac{P\sigma^3}{\epsilon}$ 
plotted against the reduced temperature for temperatures of kT = 0.5 to 3.0. Again, the red area 
is fluid, and the blue area is crystal. The general trend of the values for the transition 
temperatures and pressures is what would be expected, with the WCA fluid becoming a solid at
a higher pressure for a higher temperatuer, or becoming a liquid at a higher temperature 
for a higher pressure. But the transition pressures and temperatures are consistently higher 
than the Monte-Carlo simulations shown by the white line, and the descripancy grows as 
the temperature increases.

\begin{figure}
  \centering
  %\includegraphics[width=\columnwidth]{figs/Phase_Diagram_of_T_vs_n}
  \includegraphics[height=10cm]{figs/Phase_Diagram_of_T_vs_n}
  \caption{Phase Diagram of the reduced temperature verses the reduced density for the WCA fuild. 
  White lines show the liquid and solid transition densities at each temperature from 
  Monte-Carlo simulations of the WCA fluid. The gray region is the region of coexistence.}
  \label{fig:Phase_Diagram_of_T_vs_n}
\end{figure}

\begin{figure}
  \centering
  %\includegraphics[width=\columnwidth]{figs/Phase_Diagram_of_P_vs_T}
  \includegraphics[height=10cm]{figs/Phase_Diagram_of_P_vs_T}
  \caption{Phase Diagram of the reduced pressure verses the reduced temperature. 
  The green line marks the boundary from Monte-Carlo simulations.}
  \label{fig:Phase_Diagram_P_vs_T}
\end{figure}

%Plots of

\begin{figure}
  \centering
  %\includegraphics[width=\columnwidth]{figs/T-vs-n_at_fixed_P}
  \caption{Plot of the reduced temperature verses the reduced density. 
  The grey region indicates coexistence.}
  \label{fig:T-vs-n_at_fixed_P}
\end{figure}

\begin{figure}
  \centering
  %\includegraphics[width=\columnwidth]{figs/p-vs-T_at_fixed_n}
  \caption{Plot of the reduced pressure verses the reduced temperature. 
  The lines bending downward indicate freezing.}
  \label{fig:p-vs-T_at_fixed_density}
\end{figure}

\begin{figure}
  \centering
  %\includegraphics[width=\columnwidth]{figs/p-vs-V_at_fixed_T}
  \caption{Plot of the reduced pressure verses the reduced volume. 
  The grey region indicates coexistence.}
  \label{fig:p-vs-V_at_fixed_T}
\end{figure}

\begin{figure}
  \centering
  %\includegraphics[width=\columnwidth]{figs/p-vs-n_at_fixed_T}
  \caption{Plot of the reduced pressure verses the reduced density. 
  The grey region indicates coexistence.}
  \label{fig:p-vs-n_at_fixed_T}
\end{figure}

@@

  %\cite{TschoppBHextended}

	%1-21	
	%\cite{lucretius}
%\cite{rowlinson1964statistical}
%\cite{gil-villegas-1997-SAFT-VR, clark2006developing,
%  lafitte2013accurate}
%\cite{carnahan1969equation}
%\cite{rosenfeld1989, rosenfeld1997,
%  roth2002whitebear}
%\cite{rosenfeld1989}
	%\cite{cuesta1997dimensional, hansen2009fundamental, marechal2013density, hughes2013classical, krebs2014improved}
%\cite{cuesta1997dimensional, hansen2009fundamental, marechal2013density}
	%\cite{schmidt1999density}
%\cite{schmidt2000density, groh2001density, kim2001adsorption,
%  sweatman2002fundamental}
	%\cite{schmidt2000fluid, sweatman2002fundamental}
%\cite{weeks1971}
	%\cite{rosenfeld1989}
	%\cite{schmidt1999density}
	%%\cite{tarazonaFIXME,santos2012phi3}
	%\cite{schmidt1999density} 
	%\cite{schmidt2000fluid}
	%\cite{schmidt1999density, schmidt2000fluid}
	%\cite{schmidt1999density}
	%\citenum{schmidt2000fluid}
	%\citenum{schmidt2000fluid}
	%a Face Centered Cubic (FCC) lattice is formed~\cite{reference!}
%\cite{barker1967perturbation}
%\cite{roth2002whitebear} 
	%\cite{roth2002whitebear}
%\cite{mikolaj2004structure, eggert2002quantitative,
%yarnell1973structure}
	%\cite{verlet1967computer}
